Unlocking Sensitivity for Visibility-based Estimators of the 21\,cm Reionization Power Spectrum

### 2017-10-13 Rotation paper/draft_submission1.tex
addressed referee comments 1,2,3,6,7

With implicit bounds of integrals from $-\infty$ to $\infty$, we have:
\begin{eqnarray}
 &&\langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)\rangle \nonumber \\
 && =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\langle T^{*}(\boldsymbol{r})T(\boldsymbol{r'})\rangle B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}) \nonumber \\
 &&  =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\boldsymbol{\kappa})T(\boldsymbol{\kappa'})\rangle e^{-i(\boldsymbol{\kappa}\cdot \boldsymbol{r}-\boldsymbol{\kappa'}\cdot\boldsymbol{r'})}\right) \nonumber \\
 && \qquad  \times B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\nonumber \\
 && =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\boldsymbol{\kappa}\cdot(\boldsymbol{r}-\boldsymbol{r'})}\right) \nonumber \\
 && \qquad \times B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}) \nonumber \\
 && \approx P(k_{\bar{b}, \tau})\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2} \delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'}) \Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}) \nonumber \\
 && = P(k_{\bar{b}, \tau})\int \frac{d^{3}r}{(X^2Y)^2}B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r}) e^{-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma} \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}}, \qquad
\label{eq:main}
\end{eqnarray}
where $\delta_D$ is the Dirac delta-function and
\begin{eqnarray}
\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\equiv &\exp\left[i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\boldsymbol{\Gamma}\hat{\boldsymbol{r}}'\right)\right] \nonumber \\
& \times \exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)-i\psi_{\nu}\right].
\end{eqnarray}
%
%\begin{widetext}
%\begin{equation}
%\begin{aligned} \langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)\rangle
% & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\langle T^{*}(\boldsymbol{r})T(\boldsymbol{r'})\rangle B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\\
% & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\boldsymbol{\kappa})T(\boldsymbol{\kappa'})\rangle e^{-i(\boldsymbol{\kappa}\cdot \boldsymbol{r}-\boldsymbol{\kappa'}\cdot\boldsymbol{r'})}\right)B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\\
% & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\boldsymbol{\kappa}\cdot(\boldsymbol{r}-\boldsymbol{r'})}\right)B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\\
% & \approx P(k_{\bar{b}, \tau})\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2} \delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'}) \Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\\
% & = P(k_{\bar{b}, \tau})\int \frac{d^{3}r}{(X^2Y)^2}B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r})\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma} \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right],
%\end{aligned}
%\label{eq:main}
%\end{equation}
%
%where $\delta_D$ is the Dirac delta-function and
%\begin{equation}
%\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\equiv\exp\left[i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\boldsymbol{\Gamma}\hat{\boldsymbol{r}}'\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)-i\psi_{\nu}\right].
%\end{equation}
%
%\end{widetext}
The third equality of Eq.(\ref{eq:main}) follows from assuming the translational invariance of the statistics of the 21\,cm field. 
For the inequality in the fourth line, we have assumed that the  3D power spectrum varies negligibly over the $k$-space of interest and have thus pulled out of the integral the power spectrum centered at 
\[
k_{\boldsymbol{b}, \tau} \equiv 2\pi \sqrt{\left(\frac{\tau}{Y}\right)^2 + \left(\frac{\bar{b}}{\lambda X}\right)^2},
\label{eq:kbtau}
\]
where $\bar{b}$ is approximately the mean of $b$ and $b'$. 
In Section \ref{sec:chromaticity} we will examine the exact form of the Fourier-space footprint that is being probed, as well as its effect on foreground isolation.  
%Since $\boldsymbol{\Gamma}$ is a sky rotation, it does not affect $\nu$, hence we have taken $\nu_{r}$ outside the parenthesis in the phase term $\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma} \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)\right]$. 
Notice that the phase factor $\exp\left[-i2\pi\tau\left(\nu-\nu'\right)\right]$
drops out in the end.

@@

@@

\begin{figure}[H]
\includegraphics[width=1\linewidth]{phi_nu}

@@

\begin{figure}[H]
\includegraphics[width=1.1\linewidth]{rephs_rb}

>>>>

With implicit bounds of integrals from $-\infty$ to $\infty$, we have
\begin{eqnarray}
 &&\langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)\rangle \nonumber \\
 && =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\langle T^{*}(\boldsymbol{r})T(\boldsymbol{r'})\rangle B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}) \nonumber \\
 &&  =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\boldsymbol{\kappa})T(\boldsymbol{\kappa'})\rangle e^{-i(\boldsymbol{\kappa}\cdot \boldsymbol{r}-\boldsymbol{\kappa'}\cdot\boldsymbol{r'})}\right) \nonumber \\
 && \qquad  \times B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\nonumber \\
 && =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\boldsymbol{\kappa}\cdot(\boldsymbol{r}-\boldsymbol{r'})}\right) \nonumber \\
 && \qquad \times B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}),
 \end{eqnarray}
 where 
 \begin{eqnarray}
\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\equiv &\exp\left[i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\boldsymbol{\Gamma}\hat{\boldsymbol{r}}'\right)\right] \nonumber \\
& \times \exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)-i\psi_{\nu}\right].
\end{eqnarray}
The third equality of Eq.(\ref{eq:main}) follows from assuming the translational invariance of the statistics of the 21\,cm field. Continuing, we may make the assumption that the 3D power spectrum varies negligibly over the $k$-space of interest, thus allowing us to pull out of the integral the power spectrum centered at 
\[
k_{\boldsymbol{b}, \tau} \equiv 2\pi \sqrt{\left(\frac{\tau}{Y}\right)^2 + \left(\frac{\bar{b}}{\lambda X}\right)^2},
\label{eq:kbtau}
\]
where $\bar{b}$ is approximately the mean of $b$ and $b'$. This gives 
 \begin{eqnarray}
  &&\langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)\rangle \nonumber \\
 && \approx P(k_{\bar{b}, \tau})\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2} \delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'}) \Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}) \nonumber \\
 && = P(k_{\bar{b}, \tau})\!\!\int\!\! \frac{d^{3}r}{(X^2Y)^2}B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r}) e^{-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma} \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}}, \qquad
\label{eq:main}
\end{eqnarray}
where $\delta_D$ is the Dirac delta-function, and notice that the phase factor $\exp\left[-i2\pi\tau\left(\nu-\nu'\right)\right]$
drops out in the end. In factoring out the power spectrum from the integral, one is essentially making the approximation that the estimator probes only a single Fourier mode. In Section \ref{sec:chromaticity} we will relax this assumption and examine the exact form of the Fourier-space footprint that is being probed, as well as its effect on foreground isolation.

%
%\begin{widetext}
%\begin{equation}
%\begin{aligned} \langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)\rangle
% & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\langle T^{*}(\boldsymbol{r})T(\boldsymbol{r'})\rangle B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\\
% & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\boldsymbol{\kappa})T(\boldsymbol{\kappa'})\rangle e^{-i(\boldsymbol{\kappa}\cdot \boldsymbol{r}-\boldsymbol{\kappa'}\cdot\boldsymbol{r'})}\right)B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\\
% & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\boldsymbol{\kappa}\cdot(\boldsymbol{r}-\boldsymbol{r'})}\right)B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\\
% & \approx P(k_{\bar{b}, \tau})\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2} \delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'}) \Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\\
% & = P(k_{\bar{b}, \tau})\int \frac{d^{3}r}{(X^2Y)^2}B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r})\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma} \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right],
%\end{aligned}
%\label{eq:main}
%\end{equation}
%
%where $\delta_D$ is the Dirac delta-function and
%\begin{equation}
%\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\equiv\exp\left[i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\boldsymbol{\Gamma}\hat{\boldsymbol{r}}'\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)-i\psi_{\nu}\right].
%\end{equation}
%
%\end{widetext}
%Since $\boldsymbol{\Gamma}$ is a sky rotation, it does not affect $\nu$, hence we have taken $\nu_{r}$ outside the parenthesis in the phase term $\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma} \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)\right]$.

@@

\begin{figure}[t]
\includegraphics[width=1\linewidth]{phi_nu}

@@

\begin{figure}[t]
\includegraphics[width=1.1\linewidth]{rephs_rb}

@@
